---
title: Quarrelling with Hugo
author: ~
date: '2018-08-14'
slug: firstpost
categories: 
  - R programming
  - Data visualisation
tags:
  - R
  - blogdown
  - Hugo
  - non-linear binning
  - ggplot
  - viridis
---

```{r knitr_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Blogdown is, disappointingly, not anywhere near as easy as I was led to believe.
Or possibly, Blogdown is perfectly simple and it's Hugo I have a quarrel with.
Yes, I think that's it.  I've definitely found a dodgy tutorial, because when I tried to use new_post to generate an .Rmd file, it created an .md file, which was not what I wanted at all.  RStudio seems more intelligent on that point, at least.  However, when I accidentally closed the viewer, I ended up destroying the site and re-installing it just to get the viewer to open again.  I hope there's a better way, because once there's content, that will get extremely tedious.

I want to put some R on this, so I am going to post a ghastly graphic-in-progress.
I promise, it will be less ghastly once it is actually applied to real data. This is 
actually pretty weird code for me, since it was originally designed to be a minimal 
reproducible example for a problem which I then solved.  Just about everything I write normally uses at minimum __dplyr__ and __tibble__, and is full of chains of pipes.

There is colourblindness in the family, so I am delighted to be trying out the __viridis__ package.

```{r libraries}
library(ggplot2)
# Using viridis because it is colourblind friendly
# and also easier to interpret accurately
# Need full version rather than the viridisLite which is
# in ggplot2 because we need to use the discrete option
suppressMessages(library(viridis))
```

Scientific notation is ugly in legends.

```{r setup}
# Don't want scientific notation in my output
options(scipen = 999)
```

Knock together some random data better suited to a logarithmic-style scale.  Note that I have thus far avoided dealing with NAs or empty cells.

```{r make_data}
set.seed(412875)
a <- round(runif(100, 1, 10)^seq(1, 5, by=1), 0)
df <- data.frame(
  lat = as.vector(sapply(1:10, function(x) rep(x, 10))), 
  long = seq(1, 10), count = as.integer(a))
```

My scale breaks aren't truly logarithmic, because I wanted them to be relatively
friendly to look at.

```{r prep_plot}
scale_breaks <- c(5, 10, 50, 500, 1000, 5000, 10000, 50000, 100000)
scale_labels <- c(paste0(0, "-", scale_breaks[1]), 
  sapply(seq(2, length(scale_breaks)), 
  function(x) paste0(scale_breaks[x-1], "-", scale_breaks[x])))
df$binned_count <- cut(df$count, breaks = c(0, scale_breaks), 
  labels = scale_breaks)
```

Theme inferno's fairly loud, but why not?  I'm just messing around.

```{r make_plot}
the_plot <- ggplot(df, aes(x = long, y = lat, fill = binned_count)) +
  geom_tile() +
  xlab("Longitude") +
  ylab("Latitude")
  scale_fill_viridis(name = "No. Records", 
    breaks = c(scale_breaks), 
    labels = scale_labels, 
    guide = "legend", option = "inferno", discrete = TRUE) 
```

```{r plot_the_plot, fig.width = 6, fig.height = 4, fig.cap = "Plot to demonstrate non-linear binning of spatial count data", echo = FALSE}
print(the_plot)
```

